---
title: "Final_Paper_Figures"
author: "Grace Deitzler"
date: "8/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load libraries 
library(reshape2)
library(nlme)
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(rstatix)
library(dplyr)
library(plyr)
library(smatr)
library(magrittr)
library(data.table)
library(patchwork)
library(annotate)
library(phyloseq)
library(rstatix)
library(pbapply)
library(DESeq2)
library(DT)
library(data.table)
library(vegan)
library(remotes)
celatum_colors <- c('Yellow Orange' = "#FFA630",
                    'Tea Green' = "#D7E8BA",
                    'Cadet Blue' = "#4DA1A9",
                    'Coral' = "#EF6F6C",
                    'Sage' = "#BBC385",
                    'Hyacinth' = '#94579E'
                  
                    
)
celatum_cols <- function(...) {
  cols <- c(...)
  
  if (is.null(cols))
    return (celatum_colors)
  
  celatum_colors[cols]
}

celatum_cols("Sage")

celatum_palette <- list(
  'main' = celatum_cols("Yellow Orange", "Coral", "Cadet Blue"),
  'two' = celatum_cols("Hyacinth", "Sage"),
  'three' = celatum_cols("Hyacinth", "Sage", "Tea Green")
)

celatum_pal <- function(palette = "main", reverse = FALSE, ...) {
  pal <- celatum_palette[[palette]]
  
  if (reverse) pal <- rev(pal)
  
  colorRampPalette(pal, ...)
}

scale_color_celatum <- function(palette = "main", discrete = TRUE, reverse = FALSE, ...) {
  pal <- celatum_pal(palette = palette, reverse = reverse)
  
  if (discrete) {
    discrete_scale("colour", paste0("celatum_", palette), palette = pal, ...)
  } else {
    scale_color_gradientn(colours = pal(256), ...)
  }
}

scale_fill_celatum <- function(palette = "main", discrete = TRUE, reverse = FALSE, ...) {
  pal <- celatum_pal(palette = palette, reverse = reverse)
  
  if (discrete) {
    discrete_scale("fill", paste0("celatum_", palette), palette = pal, ...)
  } else {
    scale_fill_gradientn(colours = pal(256), ...)
  }
}
```

Code for the final figures in the ASD Celatum Paper.


## Sociability: Fig 2

### Data preparation
```{r}

#read in the new file with no naclbact
soc_n <- read.csv("Sociability_Clean_no_naclbact.csv")

## removing stage 1
social <- soc_n[soc_n$Stage != "Stage 1",]

#Time
social_time <- social[colnames(social) %in% c("Time_middle", "Time_new_zone","Time_old_zone", "Stage", "Treatment", "Gender", "Distance_total")]

#separate the two stages
social_time_Stage_2<-social_time[ social_time$Stage == "Stage 2",]
social_time_Stage_3<-social_time[social_time$Stage == "Stage 3",]

colnames(social_time_Stage_2)<-c("Amendment", "Sex", "Stage", "Distance", "Center", "Stranger 1","Empty Cup")
colnames(social_time_Stage_3)<-c("Amendment", "Sex", "Stage", "Distance", "Center", "Stranger 2","Stranger 1")

#change the dataframe order 
#social_time_Stage_2<-social_time_Stage_2[,c(1,2,3,4,6,5)]
#social_time_Stage_3<-social_time_Stage_3[,c(1,2,3,4,6,5)]

#Stage 2
compare_stage2 <- list(c("Center", "Stranger 1"), c("Center", "Empty Cup"), c("Stranger 1", "Empty Cup"))

time_stranger1 <- social_time_Stage_2 %>% dplyr::select(Amendment, Sex, `Stranger 1`)
time_emptycup <- social_time_Stage_2 %>% dplyr::select(Amendment, Sex, `Empty Cup`)
time_center1 <- social_time_Stage_2 %>% dplyr::select(Amendment, Sex, `Center`)
distance <- social_time_Stage_2 %>% dplyr::select(Amendment, Sex, `Distance`)


compare_tx= list(c("PolyIC-Bact", "PolyIC-AJ"), c("NaCl-AJ", "PolyIC-AJ"), c("NaCl-AJ", "PolyIC-Bact"))
compare_sex = list(c("F", "M"))

```

## Sociability: Stranger 1

```{r}
soc_n1 <- ggplot(data= reshape2::melt(time_stranger1), aes(x= Amendment, y= value, group=Amendment)) + 
geom_boxplot(aes(fill= Amendment)) +  
xlab("") + geom_jitter() +
ylab("Time (s)") + 
facet_wrap(~Sex) +  
scale_fill_celatum(palette = "main") +  
  stat_compare_means(comparisons = compare_tx,method = "t.test", hide.ns = TRUE) +
   ggtitle("Sociability: Time spent with Stranger 1")+ 
  theme(legend.position = "none") +
  theme_bw()+
     theme(text = element_text(size=14),
axis.text.x = element_text(angle=90, hjust=1)) 
soc_n1

soc_n_all <- ggplot(data= reshape2::melt(time_stranger1), aes(x= Amendment, y= value, group=Amendment)) + 
geom_boxplot(aes(fill= Amendment)) +  
xlab("") + geom_jitter() +
ylab("Time (s)") + 
#facet_wrap(~Sex) +  
scale_fill_celatum(palette = "main") +  
  stat_compare_means(comparisons = compare_tx,method = "t.test", hide.ns = TRUE) +
   ggtitle("Sociability: Time spent with Stranger 1")+ 
  theme(legend.position = "none") +
  theme_bw()+
     theme(text = element_text(size=14),
           
           
        axis.text.x = element_text(angle=90, hjust=1)) 
soc_n_all



ggsave("sociability1_all.tiff", dpi = 800, plot = soc_n_all)
```


## Social Novelty: Stranger 2

```{r}
compare_stage3 <- list(c("Center", "Stranger 1"), c("Center", "Stranger 2"), c("Stranger 1", "Stranger 2"))

#same as above 
time_stranger2 <- social_time_Stage_3 %>% dplyr::select(Amendment, Sex, `Stranger 2`)
time_stranger1 <- social_time_Stage_3 %>% dplyr::select(Amendment, Sex, `Stranger 1`)
time_center <- social_time_Stage_3 %>% dplyr::select(Amendment, Sex, `Center`)

soc_n2 <- ggplot(data= reshape2::melt(time_stranger2), aes(x= Amendment, y= value, group=Amendment)) + 
  geom_boxplot(aes(fill= Amendment)) +  xlab("") + 
  geom_jitter() +
  ylab("Time (s)") + 
 facet_wrap (~Sex) +
stat_compare_means(comparisons = compare_tx, method= "t.test") + 
  scale_fill_celatum(palette = "main") +  theme(legend.position="none") +
  ggtitle("Social Novelty: Time with Stranger 2") +
  theme(legend.position = "none") +
  theme_bw()+
     theme(text = element_text(size=14),
        axis.text.x = element_text(angle=90, hjust=1)) 

soc_n3 <- ggplot(data= reshape2::melt(time_stranger2), aes(x= Amendment, y= value, group=Amendment)) + 
  geom_boxplot(aes(fill= Amendment)) +  xlab("") + 
  geom_jitter() +
  ylab("Time (s)") + 
 #facet_wrap (~Sex) +
stat_compare_means(comparisons = compare_tx, method= "t.test") + 
  scale_fill_celatum(palette = "main") +  theme(legend.position="none") +
  ggtitle("Social Novelty: Time with Stranger 2") +
  theme(legend.position = "none") +
  theme_bw()+
     theme(text = element_text(size=14),
        axis.text.x = element_text(angle=90, hjust=1)) 

soc_n2
#ggsave("sociability_s3_sex_new.tiff", dpi = 800, plot = soc_n2)
```



```{r}
### Treatment Specific ANOVA Stage 2
# ANOVA
time_str1_nacl <- subset(time_stranger1, Amendment == "NaCl-AJ")
#time_str1_nacl2 <- subset(time_stranger1, Amendment == "NaCl-Bact")

#time_str1_nacl <- rbind(time_str1_nacl, time_str1_nacl2)

time_str1_polyic <- subset(time_stranger1, Amendment == "PolyIC-AJ")
time_str1_polyic2 <- subset(time_stranger1, Amendment == "PolyIC-Bact")

time_str1_all <- rbind(time_str1_polyic, time_str1_polyic2,time_str1_nacl )

#interaction <- aov(`Stranger 1` ~ Sex*Amendment, data = time_str1_nacl)
#summary(interaction)

#tukey.test <- TukeyHSD(interaction)
#tukey.test

interaction2 <- aov(`Stranger 1` ~ Sex*Amendment, data = time_str1_all)
summary(interaction2)

tukey.test2 <- TukeyHSD(interaction2)
tukey.test2
```

```{r}
time_str1_nacl <- subset(time_center1, Amendment == "NaCl-AJ")
#time_str1_nacl2 <- subset(time_stranger1, Amendment == "NaCl-Bact")

time_str1_nacl <- rbind(time_str1_nacl, time_str1_nacl2)

time_str1_polyic <- subset(time_stranger1, Amendment == "PolyIC-AJ")
time_str1_polyic2 <- subset(time_stranger1, Amendment == "PolyIC-Bact")

time_str1_all <- rbind(time_str1_polyic, time_str1_polyic2,time_str1_nacl )

#interaction <- aov(`Stranger 1` ~ Sex*Amendment, data = time_str1_nacl)
#summary(interaction)

#tukey.test <- TukeyHSD(interaction)
#tukey.test

interaction2 <- aov(`Stranger 1` ~ Sex*Amendment, data = time_str1_all)
summary(interaction2)

tukey.test2 <- TukeyHSD(interaction2)
tukey.test2
```


# center time anova stage 2

```{r}
# ANOVA
time_center_nacl <- subset(time_center, Amendment == "NaCl-AJ")
#time_str1_nacl2 <- subset(time_stranger1, Amendment == "NaCl-Bact")


time_center_polyic <- subset(time_center, Amendment == "PolyIC-AJ")
time_center_polyic2 <- subset(time_center, Amendment == "PolyIC-Bact")

time_center_all <- rbind(time_center_polyic, time_center_polyic2,time_center_nacl )

#interaction <- aov(`Stranger 1` ~ Sex*Amendment, data = time_str1_nacl)
#summary(interaction)

#tukey.test <- TukeyHSD(interaction)
#tukey.test

interaction2 <- aov(`Center` ~ Sex*Amendment, data = time_center_all)
summary(interaction2)

tukey.test2 <- TukeyHSD(interaction2)
tukey.test2
```


```{r warning=FALSE, message = FALSE, echo = FALSE}
# ANOVA stage 3
time_str2_nacl <- subset(time_stranger2, Amendment == "NaCl-AJ")
#time_str2_nacl2 <- subset(time_stranger2, Amendment == "NaCl-Bact")

#time_str2_nacl <- rbind(time_str2_nacl, time_str2_nacl2)

time_str2_polyic <- subset(time_stranger2, Amendment == "PolyIC-AJ")
time_str2_polyic2 <- subset(time_stranger2, Amendment == "PolyIC-Bact")

time_str2_all <- rbind(time_str2_polyic, time_str2_polyic2,time_str2_nacl )

interaction2 <- aov(`Stranger 2` ~ Sex*Amendment, data = time_str2_all)
summary(interaction2)

tukey.test2 <- TukeyHSD(interaction2)
tukey.test2
```

## Elevated Plus Maze

Load the data and process
```{r}
epm_n<-read.csv("EPM_Clean_no_naclbact.csv")

# Get rid of groups we don't want
epm_n <- subset(epm_n, Note.Problem == "Ok")
epm_n <- subset(epm_n, !is.na(Dist))
epm_n <- subset(epm_n, TREATMENT != "GF_NaCl-AJ")

b <- table(epm_n$Sex, epm_n$TREATMENT)
b

# normalization (to level the difference in lengths of trials)
# time 
epm_n$CL_norm_time_Tot <- epm_n$CL_Total_Time / epm_n$Total_Time
epm_n$OP_norm_time_Tot <- epm_n$OP_Total_Time / epm_n$Total_Time
epm_n$CEN_norm_time_Tot <- epm_n$CEN_Total_Time / epm_n$Total_Time

epm_n$CL_norm_time_OPandCL <- epm_n$CL_Total_Time / (epm_n$CL_Total_Time + epm_n$OP_Total_Time)
epm_n$OP_norm_time_OPandCL <- epm_n$OP_Total_Time / (epm_n$CL_Total_Time + epm_n$OP_Total_Time)

# entries
epm_n$CL_norm_ent_OPandCL <- epm_n$CL_Ent / (epm_n$CL_Ent + epm_n$OP_Ent)
epm_n$OP_norm_ent_OPandCL <- epm_n$OP_Ent / (epm_n$CL_Ent + epm_n$OP_Ent)
epm_n$CEN_norm_ent_OPandCL <- epm_n$CEN_Ent / (epm_n$CL_Ent + epm_n$OP_Ent + epm_n$CEN_Ent)

# average time per entry
epm_n$CL_time_norm_by_entries <- epm_n$CL_Total_Time / epm_n$CL_Ent
epm_n$OP_time_norm_by_entries <- epm_n$OP_Total_Time / epm_n$OP_Ent
epm_n$CEN_time_norm_by_entries <- epm_n$CEN_Total_Time / epm_n$CEN_Ent

# distance normalization
epm_n$CL_norm_dist_Tot<-epm_n$CL_Dist / epm_n$Dist
epm_n$OP_norm_dist_Tot<-epm_n$OP_Dist / epm_n$Dist
epm_n$CEN_norm_dist_Tot<-epm_n$CEN_Dist / epm_n$Dist

epm_n$CL_norm_dist_OPandCL<-epm_n$CL_Dist / (epm_n$CL_Dist + epm_n$OP_Dist)
epm_n$OP_norm_dist_OPandCL<-epm_n$OP_Dist / (epm_n$CL_Dist + epm_n$OP_Dist)

epm_n_comparisons= list(c("PolyIC-Bact", "PolyIC-AJ"),c("NaCl-AJ","PolyIC-AJ"))
compare_sex= list(c("F", "M"))
```

Total active time

```{r}
interaction <- aov(Act_Time ~ TREATMENT, data = epm_n)
summary(interaction)

tukey.test <- TukeyHSD(interaction)
tukey.test

interaction <- aov(Act_Time ~ TREATMENT * Sex, data = epm_n)
summary(interaction)

tukey.test <- TukeyHSD(interaction)
tukey.test

bp2 <- ggplot(epm_n, aes(x= TREATMENT, y= CEN_Act_Time)) + 
  geom_boxplot(aes(fill=TREATMENT))+
  facet_grid(. ~ Sex)+
  xlab("Treatment") + ylab("Total Active Time")+
  geom_jitter()+
stat_compare_means(comparisons = epm_n_comparisons,method= "wilcox.test")+
   ggtitle("Total Active Time - center")+ 
  theme(legend.position = "none")+
  scale_fill_celatum() +
    theme(legend.position = "none") +
  theme_bw()+
     theme(text = element_text(size=14),
        axis.text.x = element_text(angle=90, hjust=1)) 
bp2

#ggsave("elevatedplus_new.tiff", dpi = 800, plot = bp2)

```
```{r warning=FALSE, message = FALSE, echo = FALSE}
cor3 <- ggplot(epm_n, aes(x = Dist, y = OP_Ent, color = Sex)) +
  facet_wrap(~TREATMENT)+
  geom_point(size = 3)+
  geom_smooth(se = FALSE, method = "lm") +
  xlab("Distance (cm)") + ylab("Frequency of entries (open arm)") +
   stat_cor(method = "spearman") +
  scale_color_celatum(palette = "two") +
  ggtitle("Distance traveled vs frequency of entries into open arm") +
     theme(legend.position = "none") +
  theme_classic()+
     theme(text = element_text(size=14),
        axis.text.x = element_text(angle=90, hjust=1))
  
cor3

#ggsave("elevatedpluscorr.tiff", dpi = 800, plot = cor3)
```


## Figure 4: Marble Burying

```{r}
marble <- read.csv("Marble_Clean_nonaclbact.csv")

b <- table(marble$Sex, marble$Treatment)
b

# ANOVA
interaction <- aov(PercentBuried ~ Treatment, data = marble)
summary(interaction)

tukey.test <- TukeyHSD(interaction)
tukey.test

interaction <- aov(PercentBuried ~ Treatment*Sex, data = marble)
summary(interaction)

tukey.test <- TukeyHSD(interaction)
tukey.test

# Boxplot Buried Marbles VS Treatment
comp_marble= list(c("PolyIC-Bact", "PolyIC-AJ"), c("NaCl-AJ", "PolyIC-AJ"))


p0 <- ggplot(marble, aes(x=Treatment, y=PercentBuried)) + 
   geom_boxplot(aes(fill=Treatment), 
                width = 0.5)+
  xlab("Treatment") + ylab("Percent Marbles Buried")+
  facet_grid(. ~ Sex)+
  geom_jitter()+
  stat_compare_means(comparisons = comp_marble, method = "wilcox")+
  ggtitle("Percent Buried Marbles VS Treatment")+
  scale_fill_celatum(palette = "main") +
  theme(legend.position = "none") +
  theme_bw()+
     theme(text = element_text(size=14),
           axis.text.x = element_text(angle=90, hjust=1))
p0

p1 <- ggplot(marble, aes(x=Sex, y=PercentBuried)) + 
   geom_boxplot(aes(fill=Sex), 
                width = 0.5)+
  xlab("Sex") + ylab("Percent Marbles Buried")+
  #facet_grid(. ~ Treatment)+
  geom_jitter()+
  stat_compare_means(comparisons = compare_sex, method = "wilcox")+
  scale_fill_celatum(palette = "two") +
  theme(legend.position = "none") +
  theme_bw()+
     theme(text = element_text(size=14),
           axis.text.x = element_text(angle=90, hjust=1))
p1

p1 <- ggplot(marble, aes(x=Sex, y=PercentBuried)) + 
   geom_boxplot(aes(fill=Sex), 
                width = 0.5)+
  xlab("Sex") + ylab("Percent Marbles Buried")+
  #facet_grid(. ~ Treatment)+
  geom_jitter()+
  stat_compare_means(comparisons = compare_sex, method = "wilcox")+
  scale_fill_celatum(palette = "two") +
  theme(legend.position = "none") +
  theme_bw()+
     theme(text = element_text(size=14),
           axis.text.x = element_text(angle=90, hjust=1))
p1

#gsave("marble1.tiff", dpi = 800, plot = p0)
#ggsave("marblesex.tiff", dpi = 800, plot = p1)
```

## Microbiome

### 16S: Colon

```{r}
ps.colon <- readRDS("ps_filtdeseq_colon_final_4_20_21.rds")
#subset the ps object to remove nacl bact
ps.colon.nonaclbact <- subset_samples(ps.colon, Treatment != "NaCl_Bact")

ord_colon.n<-ordinate(ps.colon.nonaclbact,method="PCoA", distance="bray")
p1 = plot_ordination(ps.colon.nonaclbact, ord_colon.n, "species", color="Genus")
p1
colon.pcoa <- plot_ordination(ps.colon.nonaclbact, ord_colon.n, color = "Treatment", title = "Colon Only: Bray PCoA | Treatment") +
  theme_set(theme_bw())+
  geom_point(size = 3) +
  scale_color_celatum() +
  stat_ellipse(type = "t") +
  theme_bw()+
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 12))

ggsave("colon_pcoa.tiff", dpi = 800, plot = colon.pcoa)

plot_ordination(ps.colon.nonaclbact, ord_colon, color = "Sex", title = "Colon Only: Bray PCoA | Sex") +
  theme_set(theme_bw())+
  geom_point(size = 3) +
  scale_color_celatum() +
  stat_ellipse(type = "t") +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 12))

```


## Duodenum and colon richness

```{r}
ps.duo <- readRDS("ps_filtdeseq_duodenum_final_4_20_21.rds")
ps.duo.nonaclbact <- subset_samples(ps.duo, Treatment != "NaCl_Bact")
otu_table(ps.duo.nonaclbact) <- otu_table(round(as((otu_table(ps.duo.nonaclbact)), "matrix")), taxa_are_rows(ps.duo.nonaclbact)) #need to coerce the OTU counts into round numbers to get R to do the measurements properly, it doesn't like decimals
df_alpha <- as.data.frame(ps.duo.nonaclbact@sam_data)
df_alpha$Simpson <- estimate_richness(ps.duo.nonaclbact, split = TRUE, measures = "Simpson")$Simpson
df_alpha$Shannon <- estimate_richness(ps.duo.nonaclbact, split = TRUE, measures = "Shannon")$Shannon

df_alpha$Observed <- estimate_richness(ps.duo.nonaclbact, split = TRUE, measures = "Observed")$Observed
duo_richness <- ggplot(data = df_alpha, aes(x = Treatment, y = Observed, fill =Treatment)) +
   geom_boxplot() +
  geom_jitter() + 
  facet_wrap(~Sex) +
  scale_color_celatum() +
  scale_fill_celatum()+
  theme_classic() +
  stat_compare_means(comparisons = my_comparisons, method= "wilcox.test", hide.ns = TRUE) +
  ggtitle("Observed Richness by sex (duodenum samples)") +
  theme(legend.position = "none") +
  theme_classic()+
     theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1)) 
#ggsave("duo_richness_new.tiff", dpi = 800, plot = duo_richness)

l1 <- ggplot(data = df_alpha, aes(x = Treatment, y = Observed, fill =Treatment)) +
  geom_jitter() + 
  geom_boxplot() +
  facet_wrap(~Sex) +
  scale_fill_celatum() +
  theme_bw() +
  stat_compare_means(comparisons = my_comparisons) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1)) +
  ggtitle("Observed Richness by sex (duodenum samples)")


l2 <- ggplot(data = df_alpha, aes(x = Treatment, y = Shannon, fill =Treatment)) +
  geom_jitter() +   geom_boxplot() +
  facet_wrap(~Sex) +
  scale_fill_celatum() +
  stat_compare_means(comparisons = my_comparisons) +
  theme_bw()+
     theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1)) +
  ggtitle("Shannon Diversity Index by sex (duodenum samples)")

l3 <- ggplot(data = df_alpha, aes(x = Treatment, y = Simpson, fill =Treatment)) +
  geom_jitter() + 
  geom_boxplot() +
  facet_wrap(~Sex) +
  theme_classic() +
  scale_fill_celatum() +
  stat_compare_means(comparisons = my_comparisons) +
   theme_bw()+
     theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1)) +
  ggtitle("Simpson Diversity Index by sex (duodenum samples)")

#ggsave("duo_observed_new.tiff", dpi = 800, plot = l1)
#ggsave("duo_simpson_new.tiff", dpi = 800, plot = l3)
#ggsave("duo_shannon.tiff", dpi = 800, plot = l2)

```

```{r}
otu_table(ps.colon.nonaclbact) <- otu_table(round(as((otu_table(ps.colon.nonaclbact)), "matrix")), taxa_are_rows(ps.colon.nonaclbact)) #need to coerce the OTU counts into round numbers to get R to do the measurements properly, it doesn't like decimals
df_alpha_colon <- as.data.frame(ps.colon.nonaclbact@sam_data)
df_alpha_colon$Observed <- estimate_richness(ps.colon.nonaclbact, split = TRUE, measures = "Observed")$Observed
df_alpha_colon$Simpson <- estimate_richness(ps.colon.nonaclbact, split = TRUE, measures = "Simpson")$Simpson
df_alpha_colon$Shannon <- estimate_richness(ps.colon.nonaclbact, split = TRUE, measures = "Shannon")$Shannon

my_comparisons <- list(c("PolyIC_Bact", "PolyIC_AJ"), c("NaCl_AJ", "PolyIC_AJ"), c("NaCl_AJ", "PolyIC_Bact"))
my_comparisons2 <- list(c("M", "F"))
p1 <-  ggplot(data = df_alpha_colon, aes(x = Treatment, y = Observed, fill =Treatment)) +
   geom_boxplot() +
  geom_jitter() +
  facet_wrap(~Sex) +
  scale_color_celatum() +
  scale_fill_celatum() +
  stat_compare_means(comparisons = my_comparisons, method= "wilcox.test", hide.ns = TRUE) +
  ggtitle("Observed Richness by sex (colon samples)") +
    theme(legend.position = "none") +
  theme_bw()+
     theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1)) 
rich_colon

rich_sex <- ggplot(data = df_alpha_colon, aes(x = Sex, y = Observed, fill = Sex)) +
  geom_jitter() + 
  geom_boxplot() +
  facet_wrap(~Treatment) +
  theme_classic() +
  scale_fill_celatum() +
  stat_compare_means(comparisons = my_comparisons2, method= "wilcox.test", hide.ns = TRUE) +
   theme(legend.position = "none") +
  theme_classic()+
     theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1)) +
  ggtitle("Observed Richness by treatment (colon samples)")
rich_sex 

p2 <- ggplot(data = df_alpha_colon, aes(x = Treatment, y = Shannon, fill =Treatment)) +
  geom_jitter() +   geom_boxplot() +
  facet_wrap(~Sex) +
  scale_fill_celatum() +
  stat_compare_means(comparisons = my_comparisons) +
  theme_bw()+
     theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1)) +
  ggtitle("Shannon Diversity Index by sex (colon samples)")

p3 <- ggplot(data = df_alpha_colon, aes(x = Treatment, y = Simpson, fill =Treatment)) +
  geom_jitter() + 
  geom_boxplot() +
  facet_wrap(~Sex) +
  theme_classic() +
  scale_fill_celatum() +
  stat_compare_means(comparisons = my_comparisons) +
   theme_bw()+
     theme(text = element_text(size=12),
        axis.text.x = element_text(angle=90, hjust=1)) +
  ggtitle("Simpson Diversity Index by sex (colon samples)")

ggsave("colon_observed_new.tiff", dpi = 800, plot = p1)
ggsave("colon_simpson_new.tiff", dpi = 800, plot = p3)
ggsave("colon_shannon.tiff", dpi = 800, plot = p2)

```


## Beta diversity
```{r}
ord_colon<-ordinate(ps.colon.nonaclbact,method="PCoA", distance="bray")

colon_bray <- plot_ordination(ps.colon.nonaclbact, ord_colon, color = "Treatment", title = "Colon Only: Bray PCoA | Treatment") +
  #theme_set(theme_bw())+
  geom_point(size = 3) +
  scale_color_celatum() +
  stat_ellipse(type = "t") +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 18))

plot_ordination(ps.colon.nonaclbact, ord_colon, color = "Sex", title = "Colon Only: Bray PCoA | Sex") +
  #theme_classic()+
  geom_point(size = 3) +
  scale_color_celatum() +
   stat_ellipse(type = "t") +
  theme(legend.position = "bottom") +
   theme(text = element_text(size = 18))

#bray_tx
ggsave("colon_bray_new.tiff", plot = colon_bray, dpi = 800)

braydist.1 <- phyloseq::distance(ps.duo.nonaclbact, "bray")
ord<-ordinate(ps.duo.nonaclbact,method="PCoA", distance="bray")


duo_bray <- plot_ordination(ps.duo.nonaclbact, ord, color = "Treatment", title = "Duodenum Only: Bray PCoA | Treatment") +
  theme_set(theme_bw())+
  geom_point(size = 3) +
  scale_color_celatum() +
  stat_ellipse(type = "t") +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 18))
ggsave("duo_bray_new.tiff", plot = duo_bray, dpi = 800)

plot_ordination(ps.duo.nonaclbact, ord, color = "Sex", title = "Duodenum Only: Bray PCoA | Sex") +
  theme_set(theme_bw())+
  geom_point(size = 3) +
  scale_color_celatum() +
  stat_ellipse(type = "t") +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 18))

```
# Splitting up the PCOA by sex: colon

```{r}
colon.ps.male <- subset_samples(ps.colon.nonaclbact, Sex=="M")
colon.ps.female <- subset_samples(ps.colon.nonaclbact, Sex=="F")

# MALES:
braydist.1.m <- phyloseq::distance(colon.ps.male, "bray")
ord.m <-ordinate(colon.ps.male,method="PCoA", distance="bray")

col.pcoa.m <- plot_ordination(colon.ps.male, ord.m, color = "Treatment", title = "Males, Colon: Bray PCoA | Treatment") +
  theme_set(theme_bw())+
  geom_point(size = 3) +
  scale_color_celatum() +
  stat_ellipse(type = "t") +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 18))



# FEMALES:
braydist.1.f <- phyloseq::distance(colon.ps.female, "bray")
ord.f <-ordinate(colon.ps.female,method="PCoA", distance="bray")

col.pcoa.f <- plot_ordination(colon.ps.female, ord.f, color = "Treatment", title = "Females, Colon: Bray PCoA | Treatment") +
  theme_set(theme_bw())+
  geom_point(size = 3) +
  scale_color_celatum() +
  stat_ellipse(type = "t") +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 18))
col.pcoa.m
col.pcoa.f

ggsave("colon_female_bray_new.tiff", plot = col.pcoa.f, dpi = 800)
ggsave("colon_male_new.tiff", plot = col.pcoa.m, dpi = 800)

```
# Splitting up the PCOA by sex: duodenum

```{r}
duo.ps.male <- subset_samples(ps.duo.nonaclbact, Sex=="M")
duo.ps.female <- subset_samples(ps.duo.nonaclbact, Sex=="F")

# MALES:
braydist.1.m <- phyloseq::distance(duo.ps.male, "bray")
ord.m <-ordinate(duo.ps.male,method="PCoA", distance="bray")

duo.pcoa.m <- plot_ordination(duo.ps.male, ord.m, color = "Treatment", title = "Males, Duodenum: Bray PCoA | Treatment") +
  theme_set(theme_bw())+
  geom_point(size = 3) +
  scale_color_celatum() +
  stat_ellipse(type = "t") +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 18))



# FEMALES:
braydist.1.f <- phyloseq::distance(duo.ps.female, "bray")
ord.f <-ordinate(duo.ps.female,method="PCoA", distance="bray")

duo.pcoa.f <- plot_ordination(duo.ps.female, ord.f, color = "Treatment", title = "Females, Duodenum: Bray PCoA | Treatment") +
  theme_set(theme_bw())+
  geom_point(size = 3) +
  scale_color_celatum() +
  stat_ellipse(type = "t") +
  theme(legend.position = "bottom") +
  theme(text = element_text(size = 18))
duo.pcoa.m
duo.pcoa.f

ggsave("duo_female_bray_new.tiff", plot = duo.pcoa.f, dpi = 800)
ggsave("duo_male_new.tiff", plot = duo.pcoa.m, dpi = 800)

```

PERMANOVA: Colon 
```{r}
factors_to_test <- c("Treatment", "Sex")

#1: extract the distance
braydist.1 <- phyloseq::distance(ps.colon.nonaclbact, "bray")

#2: make the data frame
sampledf.1 <- data.frame(sample_data(ps.colon.nonaclbact))

#3: adonis - uses the extracted distance along with the dataframe that you made to accompany the phyloseq object.
adonis(braydist.1 ~ Treatment, data = sampledf.1) 
adonis(braydist.1 ~ Sex, data = sampledf.1)
adonis(braydist.1 ~ Sex * Treatment, data = sampledf.1)

beta <- betadisper(braydist.1, sampledf.1$Sex)
permutest(beta)
beta <- betadisper(braydist.1, sampledf.1$Treatment)
permutest(beta)


#Just females

factors_to_test <- c("Treatment", "Sex")

#1: extract the distance
braydist.1 <- phyloseq::distance(colon.ps.female, "bray")

#2: make the data frame
sampledf.1 <- data.frame(sample_data(colon.ps.female))

#3: adonis - uses the extracted distance along with the dataframe that you made to accompany the phyloseq object.
adonis(braydist.1 ~ Treatment, data = sampledf.1) 
#adonis(braydist.1 ~ Sex, data = sampledf.1)
#adonis(braydist.1 ~ Sex * Treatment, data = sampledf.1)
beta <- betadisper(braydist.1, sampledf.1$Treatment)
permutest(beta)

#Just males

factors_to_test <- c("Treatment", "Sex")

#1: extract the distance
braydist.1 <- phyloseq::distance(colon.ps.male, "bray")

#2: make the data frame
sampledf.1 <- data.frame(sample_data(colon.ps.male))

#3: adonis - uses the extracted distance along with the dataframe that you made to accompany the phyloseq object.
adonis(braydist.1 ~ Treatment, data = sampledf.1) 
#adonis(braydist.1 ~ Sex, data = sampledf.1)
#adonis(braydist.1 ~ Sex * Treatment, data = sampledf.1)
beta <- betadisper(braydist.1, sampledf.1$Treatment)
permutest(beta)

```


PERMANOVA: Duodenum
```{r}
factors_to_test <- c("Treatment", "Sex")

#1: extract the distance
braydist.1 <- phyloseq::distance(ps.duo.nonaclbact, "bray")

#2: make the data frame
sampledf.1 <- data.frame(sample_data(ps.duo.nonaclbact))

#3: adonis - uses the extracted distance along with the dataframe that you made to accompany the phyloseq object.
adonis(braydist.1 ~ Treatment, data = sampledf.1) 
adonis(braydist.1 ~ Sex, data = sampledf.1)
adonis(braydist.1 ~ Sex * Treatment, data = sampledf.1)

beta <- betadisper(braydist.1, sampledf.1$Sex)
permutest(beta)
beta <- betadisper(braydist.1, sampledf.1$Treatment)
permutest(beta)

#Just females

factors_to_test <- c("Treatment", "Sex")

#1: extract the distance
braydist.1 <- phyloseq::distance(duo.ps.female, "bray")

#2: make the data frame
sampledf.1 <- data.frame(sample_data(duo.ps.female))

#3: adonis - uses the extracted distance along with the dataframe that you made to accompany the phyloseq object.
adonis(braydist.1 ~ Treatment, data = sampledf.1) 
#adonis(braydist.1 ~ Sex, data = sampledf.1)
#adonis(braydist.1 ~ Sex * Treatment, data = sampledf.1)
beta <- betadisper(braydist.1, sampledf.1$Treatment)
permutest(beta)

#Just males

factors_to_test <- c("Treatment", "Sex")

#1: extract the distance
braydist.1 <- phyloseq::distance(duo.ps.male, "bray")

#2: make the data frame
sampledf.1 <- data.frame(sample_data(duo.ps.male))

#3: adonis - uses the extracted distance along with the dataframe that you made to accompany the phyloseq object.
adonis(braydist.1 ~ Treatment, data = sampledf.1) 
#adonis(braydist.1 ~ Sex, data = sampledf.1)
#adonis(braydist.1 ~ Sex * Treatment, data = sampledf.1)
beta <- betadisper(braydist.1, sampledf.1$Treatment)
permutest(beta)
```



